\documentclass[runningheads]{llncs}

\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{enumitem}
\setlist{noitemsep,topsep=2pt,leftmargin=*}

\tikzset{
  block/.style={draw, rounded corners=2pt, align=center, minimum height=7mm,
    inner xsep=4pt, font=\scriptsize, fill=gray!8},
  smallblock/.style={draw, rounded corners=2pt, align=center, minimum height=6mm,
    inner xsep=3pt, font=\scriptsize, fill=blue!7},
  loss/.style={draw, rounded corners=2pt, align=center, minimum height=6mm,
    inner xsep=4pt, font=\scriptsize, fill=orange!12},
  arr/.style={-{Latex[length=2mm]}, line width=.35pt}
}

\begin{document}

\title{Siamese-based Real-ESRGAN with High-Order Degradation for Realistic Image Restoration}
\titlerunning{Siamese-based Real-ESRGAN for Image Restoration}

\author{Le Hoai Nam\inst{1} \and Doan Nhat Quang\inst{1} \and Giang Anh Tuan\inst{1} \and
Tran Hoang Tung\inst{1} \and Le Huu Ton\inst{2} \and Nguyen Hoang Ha\inst{1}}
\authorrunning{L. H. Nam et al.}

\institute{University of Science and Technology of Hanoi, Vietnam Academy of Science and Technology, Hanoi, Vietnam\\
\email{nguyen-hoang.ha@usth.edu.vn} \and
CMC University, Hanoi, Vietnam}

\maketitle

\begin{abstract}
Real-world image restoration remains challenging because practical degradations combine blur, sensor noise, chromatic aberration, and JPEG compression in ways that are difficult to reproduce synthetically. We present a Siamese-based Real-ESRGAN framework that jointly improves degradation diversity and training robustness. First, we extend the high-order degradation pipeline with randomized operation ordering, refined compression ranges, explicit camera-noise simulation, and four offline quality levels per image (90\%, 80\%, 70\%, and 60\%). Second, we introduce a weight-sharing dual-branch training strategy: a Teacher pass processes the milder degradation while a Student pass restores the harder degradation, with output and feature consistency losses linking both passes. A three-phase curriculum progressively increases the degradation gap. Experiments on standard and real-world benchmarks show that the proposed Siamese model improves perceptual quality, achieving NIQE 3.82 on Urban100 and an 18.7\% NIQE gain over Real-ESRGAN on RealSR\_V3 Canon, while keeping inference cost identical to the single-branch baseline.

\keywords{Image Super-Resolution \and Siamese Network \and Knowledge Distillation \and Realistic Degradation \and Real-ESRGAN}
\end{abstract}

\section{Introduction}

Single-image super-resolution (SR) aims to recover high-quality images from low-quality observations. Deep SR methods have achieved strong performance under synthetic bicubic degradation, but real photographs usually contain mixed and unknown artifacts from optics, sensors, compression, and post-processing. GAN-based methods such as ESRGAN~\cite{wang2018esrgan} and BSRGAN~\cite{zhang2021designing} improve perceptual sharpness, and Real-ESRGAN~\cite{wang2021realesrgan} further improves blind SR by synthesizing high-order degradations. However, training on a single degradation distribution can still be unstable when the target images contain stronger or differently ordered artifacts.

This work extends our previous Real-ESRGAN refinement study~\cite{11309534}. The earlier work focused on improved degradation synthesis and conventional knowledge distillation for model compression. Here, we address a different problem: robust restoration under gradually increasing degradation severity. Instead of using a fixed large Teacher and a separate Student, we use one shared-weight generator in two passes. The cleaner input acts as dynamic guidance for the harder input, and the degradation gap is increased through curriculum learning.

Our contributions are:
\begin{itemize}
    \item A weight-sharing Siamese Real-ESRGAN training scheme that performs Teacher and Student passes through the same RRDBNet generator, adding no inference-time parameters.
    \item A three-phase curriculum that progresses from 90\%/80\% to 80\%/70\% and 70\%/60\% degradation pairs, improving stability under severe artifacts.
    \item An enhanced high-order degradation pipeline with randomized operation order, realistic compression ranges, explicit camera noise, and multi-level offline generation.
\end{itemize}

\section{Related Work}

\noindent\textbf{Real-world super-resolution.}
SRCNN~\cite{dong2014learning} introduced deep learning to SR, followed by stronger residual architectures such as EDSR~\cite{lim2017enhanced} and RDN~\cite{zhang2018residualdensenetworkimage}. These methods usually assume bicubic degradation. Blind SR methods~\cite{zhang2018learning,luo2020unfolding,gu2019blind} relax this assumption by estimating or modeling unknown degradations. Real-ESRGAN~\cite{wang2021realesrgan} uses high-order synthetic degradation to approximate real camera artifacts; our method keeps this direction but generates multiple controlled degradation levels for curriculum training.

\noindent\textbf{Knowledge distillation and Siamese learning.}
Knowledge Distillation (KD)~\cite{hinton2015distilling,Hui_2019} is widely used to compress restoration networks, and feature-aware distillation methods such as FAKD~\cite{gao2022featuredistillationinteractionweighting} preserve structural details. Conventional KD relies on a fixed Teacher. Siamese networks~\cite{bromley1993signature,koch2015siamese}, in contrast, learn from paired inputs under a shared representation. We combine these ideas by using two degradation levels of the same image as a self-guided pair.

\noindent\textbf{Curriculum and generative SR.}
Curriculum learning orders examples from easy to hard to improve convergence~\cite{bengio2009curriculum,soviany2022curriculum}. We define difficulty by degradation strength. Recent diffusion-based SR models such as StableSR~\cite{wang2024exploiting}, DiffBIR~\cite{lin2023diffbir}, and PASD~\cite{yang2023pixel} produce high perceptual fidelity but are computationally expensive. Our GAN-based framework targets practical inference speed while improving realism.

\section{Methodology}

The proposed framework has two components: multi-level realistic degradation and weight-sharing Siamese training. Fig.~\ref{fig:overview} summarizes the full pipeline.

\begin{figure}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[node distance=4.5mm and 7mm]
  \node[block] (gt) {GT image};
  \node[block, right=of gt] (crop) {$480{\times}480$\\patches};
  \node[block, right=of crop] (deg) {High-order\\degradation};
  \node[smallblock, above right=2mm and 8mm of deg] (q90) {90\%};
  \node[smallblock, right=of deg] (q80) {80\%};
  \node[smallblock, below right=2mm and 8mm of deg] (q70) {70\% / 60\%};
  \node[block, right=18mm of q80] (pair) {Paired inputs\\same GT};
  \node[block, right=of pair] (teacher) {Teacher pass\\milder input};
  \node[block, below=5mm of teacher] (student) {Student pass\\harder input};
  \node[loss, right=of teacher] (loss) {Pixel, perceptual, GAN\\output KD, feature KD};
  \node[block, right=of loss] (sr) {Restored image};

  \draw[arr] (gt) -- (crop);
  \draw[arr] (crop) -- (deg);
  \draw[arr] (deg) -- (q90);
  \draw[arr] (deg) -- (q80);
  \draw[arr] (deg) -- (q70);
  \draw[arr] (q90.east) -- (pair.west);
  \draw[arr] (q80.east) -- (pair.west);
  \draw[arr] (q70.east) -- (pair.west);
  \draw[arr] (pair) -- (teacher);
  \draw[arr] (pair) -- (student);
  \draw[arr] (teacher) -- (loss);
  \draw[arr] (student) -- (loss);
  \draw[arr] (loss) -- (sr);
  \node[draw, dashed, rounded corners=2pt, fit=(teacher)(student), inner sep=3pt,
        label={[font=\scriptsize]below:shared RRDBNet weights}] {};
\end{tikzpicture}}
\caption{Overview of the proposed Siamese Real-ESRGAN framework. Multi-level degraded samples are paired by curriculum stage and processed by one shared-weight generator in two passes.}
\label{fig:overview}
\end{figure}

\subsection{Multi-Level High-Order Degradation}

Real-ESRGAN synthesizes low-quality images through repeated resizing, blurring, noise injection, and JPEG compression. We retain the two-stage structure but introduce three refinements. First, after resizing, blur, noise, and compression are randomly ordered, so the model sees more realistic artifact combinations. Second, JPEG quality is limited to practical ranges to avoid synthetic blocking patterns that rarely appear in the target data. Third, camera-specific noise, including chromatic aberration and sensor shot/read noise, is explicitly simulated.

For $\times4$ SR, GT images are cropped into $480\times480$ patches with 50\% overlap. Their LQ counterparts are generated at $120\times120$. Each GT patch is pre-degraded into four quality levels: 90\%, 80\%, 70\%, and 60\%, where lower values correspond to stronger composite degradation. Offline generation improves reproducibility and allows efficient pairing during training.

\subsection{Siamese Training with Shared Weights}

Our generator follows the Real-ESRGAN RRDBNet backbone~\cite{wang2018esrgan,wang2021realesrgan} with 23 RRDB blocks and 64 feature channels. During training, each mini-batch contains two degraded versions of the same GT image. The Teacher pass receives the milder degradation and runs under \texttt{torch.no\_grad()}, producing $T_{HQ}$ and intermediate feature maps $T_{feat}$. The Student pass receives the harder degradation and produces $S_{HQ}$ and $S_{feat}$. Both passes use identical generator weights; only the Student pass contributes gradients.

The training objective combines reconstruction, perceptual, adversarial, and consistency losses:
\[
L_{total}=w_{pix}L_{pix}+w_{per}L_{per}+w_{GAN}L_{GAN}
\;+\lambda^{out}_{kd}L^{out}_{kd}+\lambda^{feat}_{kd}L^{feat}_{kd}.
\]
$L_{pix}$ is the L1 loss between $S_{HQ}$ and GT, $L_{per}$ is a VGG19 perceptual loss~\cite{simonyan2014very,johnson2016perceptual}, and $L_{GAN}$ is computed with a U-Net discriminator using spectral normalization. $L^{out}_{kd}$ aligns $S_{HQ}$ with $T_{HQ}$, while $L^{feat}_{kd}$ aligns $S_{feat}$ with $T_{feat}$. Since the Teacher output is detached, it acts as a dynamic pseudo-target without requiring a second stored network.

\subsection{Curriculum Schedule}

Table~\ref{tab:curriculum} shows the three training phases. Starting from mild pairs stabilizes GAN training; later phases gradually increase the restoration difficulty.

\begin{table}[t]
\centering
\caption{Three-phase degradation curriculum. Quality 100\% denotes the GT image; lower values indicate stronger degradation.}
\label{tab:curriculum}
\small
\begin{tabular}{@{}cccc@{}}
\toprule
\textbf{Phase} & \textbf{Teacher input} & \textbf{Student input} & \textbf{Focus} \\
\midrule
1 & 90\% & 80\% & Stable warm-up \\
2 & 80\% & 70\% & Best perceptual balance \\
3 & 70\% & 60\% & Severe degradation robustness \\
\bottomrule
\end{tabular}
\end{table}

\section{Experiments}

\subsection{Experimental Setup}

Training uses the same sources as our previous work~\cite{11309534}: DF2K (DIV2K, Flickr2K, and OutdoorScene), Unsplash, and RealSR. Only HQ images are retained; all LQ images are generated by our degradation pipeline. Evaluation is conducted on Set5, Set14, RealSR\_V3 Canon/Nikon, BSDS100, and Urban100.

We report PSNR, SSIM, and NIQE~\cite{mittal2012making}. PSNR and SSIM measure distortion fidelity, while NIQE is a no-reference naturalness metric where lower values indicate better perceptual quality. For real-world SR, NIQE often better reflects visual realism than pixel-level metrics~\cite{wang2021realesrgan}.

Training follows Real-ESRGAN settings with Siamese modifications. We use Adam ($\beta_1=0.9$, $\beta_2=0.999$), initial learning rate $1\times10^{-4}$, MultiStepLR decay ($\gamma=0.5$) at 200k, 400k, 500k, and 700k iterations, and 800k total iterations with 10k warm-up. Loss weights are $w_{pix}=2.0$, $w_{GAN}=0.5$, $\lambda^{out}_{kd}=\lambda^{feat}_{kd}=0.15$, and a consistency term of 0.075 after warm-up. Experiments run on an NVIDIA RTX 3090. Because both branches share one 16.7M-parameter backbone, inference cost remains the same as Real-ESRGAN, about 0.08 s for a $720\times480$ image.

\subsection{Quantitative Results}

Table~\ref{tab:quantitative_results} compares the proposed method with representative baselines. The Siamese Phase 2 model achieves the best NIQE on Urban100 (3.82) and strong real-world perceptual quality. On RealSR\_V3 Canon, it improves NIQE from 7.02 for Real-ESRGAN to 5.71, an 18.7\% reduction. Phase 3 further improves NIQE on RealSR Canon/Nikon and Set14, but sacrifices more PSNR/SSIM due to the heavier degradation target.

\begin{table*}[t]
\centering
\caption{Comparison on benchmark datasets for $\times4$ SR. Best NIQE per dataset is in \textbf{bold}.}
\label{tab:quantitative_results}
\renewcommand{\arraystretch}{1.05}
\resizebox{\textwidth}{!}{%
\begin{tabular}{llccccccccccccc}
\toprule
\multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\makecell{\textbf{Training}\\\textbf{Data}}}
  & \multicolumn{2}{c}{\textbf{Set5}}
  & \multicolumn{2}{c}{\textbf{Set14}}
  & \multicolumn{2}{c}{\textbf{RealSR Canon}}
  & \multicolumn{2}{c}{\textbf{RealSR Nikon}}
  & \multicolumn{2}{c}{\textbf{BSDS100}}
  & \multicolumn{2}{c}{\textbf{Urban100}}
  & \multirow{2}{*}{\makecell{\textbf{Time}\\\textbf{(s)}}} \\
\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}\cmidrule(lr){11-12}\cmidrule(lr){13-14}
 & & \textbf{PSNR/SSIM} & \textbf{NIQE}
   & \textbf{PSNR/SSIM} & \textbf{NIQE}
   & \textbf{PSNR/SSIM} & \textbf{NIQE}
   & \textbf{PSNR/SSIM} & \textbf{NIQE}
   & \textbf{PSNR/SSIM} & \textbf{NIQE}
   & \textbf{PSNR/SSIM} & \textbf{NIQE} & \\
\midrule
RealSR-JPEG~\cite{ji2020realworld}
  & Bicubic & 24.74/0.698 & 6.11 & 23.38/0.602 & 6.01
  & 25.66/0.720 & 7.57 & 25.49/0.698 & 6.68
  & 19.21/0.440 & 4.37 & 17.66/0.490 & 3.97 & 1.48 \\
BSRGAN~\cite{zhang2021designing}
  & High-order & 24.55/0.706 & 6.71 & 23.42/0.624 & 5.72
  & 25.02/0.748 & \textbf{5.37} & 24.41/0.718 & 5.68
  & 21.96/0.551 & 4.51 & 19.48/0.575 & 4.08 & 1.42 \\
SwinIR~\cite{liang2021swinir}
  & High-order & \textbf{30.72}/\textbf{0.872} & 7.28 & \textbf{27.01}/\textbf{0.754} & 6.29
  & 25.88/0.749 & 9.08 & 25.71/0.731 & 8.61
  & 19.91/0.563 & 6.81 & 16.65/0.529 & 5.31 & 2.72 \\
Real-ESRGAN~\cite{wang2021realesrgan}
  & Original & 26.31/0.801 & 7.89 & 25.18/0.691 & 5.88
  & 25.84/0.780 & 7.02 & 25.19/0.755 & 6.69
  & \textbf{22.34}/0.568 & 4.46 & \textbf{20.12}/\textbf{0.639} & 4.33 & 1.55 \\
\makecell[l]{\textbf{Ours -- Student}}
  & Our Data & 25.19/0.767 & \textbf{6.54} & 23.89/0.694 & 5.82
  & \textbf{26.77}/\textbf{0.805} & 6.78 & \textbf{25.93}/\textbf{0.763} & 6.25
  & 22.31/0.572 & 4.82 & 20.00/0.598 & 4.05 & 1.45 \\
\makecell[l]{\textbf{Ours -- Siamese Ph.2}}
  & Our Data & 23.37/0.718 & 7.98 & 22.69/0.642 & 5.91
  & 23.21/0.748 & 5.71 & 22.76/0.713 & 5.69
  & 21.45/\textbf{0.573} & \textbf{4.25} & 18.57/0.586 & \textbf{3.82} & 1.72 \\
\makecell[l]{\textbf{Ours -- Siamese Ph.3}}
  & Our Data & 21.98/0.658 & 8.00 & 21.44/0.590 & \textbf{5.58}
  & 22.76/0.733 & \textbf{5.64} & 22.28/0.695 & \textbf{5.47}
  & 20.70/0.542 & 4.28 & 17.80/0.552 & 3.93 & \textbf{1.37} \\
\bottomrule
\end{tabular}}
\end{table*}

The reduced PSNR/SSIM of the Siamese phases on synthetic benchmarks reflects the perception-distortion trade-off~\cite{blau2018perception}. The model prioritizes natural textures and artifact suppression over exact pixel reconstruction, which is preferable for blind real-world restoration.

\subsection{Ablation and Qualitative Analysis}

Training directly on the hardest 70\%/60\% pair without curriculum caused unstable GAN behavior and noisy textures during early iterations. The progressive schedule avoided this failure mode. Replacing the shared-weight Teacher with a separate fixed Teacher increased memory use without improving NIQE, indicating that dynamic same-backbone guidance is sufficient. Removing feature-level KD degraded texture consistency, especially around repeated structures.

Qualitatively, the Siamese Phase 2 model produces sharper text and cleaner edges on RealSR\_V3 Canon, reduces aliasing on Urban100 building patterns, and preserves fine BSDS100 textures with fewer hallucinated artifacts than Real-ESRGAN and SwinIR. These observations match the NIQE improvements in Table~\ref{tab:quantitative_results}.

\section{Conclusion}

We proposed a Siamese Real-ESRGAN framework for realistic image restoration. The method combines multi-level high-order degradation, a shared-weight Teacher--Student training pass, and a three-phase curriculum. It improves perceptual quality on real-world and structural benchmarks while preserving the inference cost of the original single-branch Real-ESRGAN. Future work will explore automatic curriculum scheduling and lightweight generative refinement for more severe degradations.

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
